---
title: "Simple Linear Programming Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Business Problem

XYZ computer company produces laptop and desktop computers. The expected demand for laptops to be at least 1,000, and for desktops to be at least 800 per day. (Assumption: production is in accordance with the demand) The production facility has a limited capacity of no more than 2,000 laptops and 1,700 desktops per day. At most 2,000 computers per day must be satisfied. Each laptop computer generates $600 net profit and each desktop computer generates $300 net profit.

The company wants to determine how many of each type should be made daily to maximize net profits.

## Solution

This is an Optimization (a type of prescriptive modeling) problem. Since the objective is to maximize (optimize) profit, optimization using Linear Programming (LP) may be used to solve this problem.

###Problem Summary:

|                    | Laptop     | Desktop  | Total   |
| ------------------ |:----------:| --------:| -------:|
| Demand/day         | >=1000     | >=800    |         |
| Capacity/day       | <=2000     | <=1700   |         |
| Profit($)/computer | 600        |   300    | Maximize|
| Computers          |    X       |   Y      | <=2000  |
|                    |            |          |         |


####Decision Variables:
**X**: number of laptops made daily

**Y**: number of desktops made daily



####Objective Function:
**600*X + 300*Y = Z (Maximum Profit)**



####Constraints:
i) Total production of laptop and desktop per day must be less than or equal to 2000: **X+Y <=2000**

ii) Production facility has the limitation of maximum 2000 laptops: **X<=2000**

iii) Production facility has the limitation of maximum 1700 desktops: **Y<=1700**

iv) The demand for laptops is at least 1000: **X>=1000**

v) The demand for desktops is at least 800 respectively: **Y>=800**



```{r}
# install lpSolveAPI
if(!require(lpSolveAPI)){install.packages("lpSolveAPI")}

# use libray
library(lpSolveAPI)

# Develop model named lpcom with 5 constraints and 2 decision variables
lpcom <- make.lp(5,2)

# develop objective function to maximize profit
lp.control(lpcom, sense="max")
set.objfn(lpcom, c(600,300))

# add constraints to the model
add.constraint(lpcom, c(1,1),"<=", 2000)
add.constraint(lpcom, c(1,0), ">=", 1000)
add.constraint(lpcom, c(1,0), "<=", 2000)
add.constraint(lpcom, c(0,1), ">=", 800)
add.constraint(lpcom, c(0,1), "<=", 1700)

# view the model
lpcom

# solve the model
solve(lpcom)

# get objective value (maximum profit)
get.objective(lpcom)

# get the values of decision variables
get.variables(lpcom)

```


##Results:
1. Optimal number of laptop is 1200 and desktop is 800
2. The value of the objective function (total net profit) is : $960,000

